---
title: "Looking at Police Misconduct Incidents in NYC"
author: "Becca Tramel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Looking at the Data

The file *locationsstrat2* was built using [FiveThirtyEight's data on police misconduct settlements](https://github.com/fivethirtyeight/police-settlements). It contains 100 random cases from each of  the years from 2008 - 2018. These years were chosen because each has at least 1000 cases in the original data set. 

We then cleaned the *location* column and used the package *ggmap* to obtain the latitude and longitude of each incident. 

```{r}
library(tidyverse)

locationstrat2 <- read_csv(url("https://raw.githubusercontent.com/becca-t/policemisconductnyc/main/locationsstrat2"))

```

## Using ggmap() to Create a Map

We can use [ggmap](http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf) to create a map of NYC. Then we'll use *ggplot2* to plot our locations on top of this map. At first we'll plot them together, then break them up by year.

I found [this blog](https://www.jessesadler.com/post/geocoding-with-r/) incredibly helpful in learning about some of the features of the ggmap package. 

First we'll enter our key for Google's API.

```{r, echo = FALSE}
library(ggmap)

register_google(key = "")
```

Then we'll create our map of NYC. 

```{r, message=FALSE}
ny_map <- get_googlemap(center = c(lon = -74.0060, lat = 40.7128), zoom = 11)
```

```{r}
ggmap(ny_map)
```

Now we can use *ggplot2* to add our data to the map. I'm going to make two versions below. One will include all of the data in one map, and the other will create a map for each year. 

```{r}
# Making a single map:
ggmap(ny_map) + 
  geom_point(data = locationstrat2, aes(x = lon, y = lat), alpha = 0.2)

# Making a map for each year:
ggmap(ny_map) + 
  geom_point(data = locationstrat2, aes(x = lon, y = lat), alpha = 0.2) +
  facet_wrap(~incident_year)

```


